---
title: "Using `nullranges::matchranges()` with `BentoBox`"
author: "Eric S. Davis"
date: "`r format(Sys.Date(), '%m/%d/%Y')`"
output:
  rmarkdown::html_document:
    highlight: tango
    toc: true
    toc_float: true	
    fig_width: 5
    fig_height: 3
vignette: |
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Using nullranges::matchranges() with BentoBox}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
knitr::opts_knit$set(global.device = TRUE)
```

This lightning talk will attempt to use a toy dataset to generate covariate-matched, null-hypothesis GRanges with `nullranges::matchRanges()` and then visualize the results using the `BentoBox` framework.

## Generating covariate-matched GRanges

For this example lets use a utility function to create a simulated dataset:

```{r, message=FALSE, warning=FALSE}
set.seed(123)
x <- Bioc2021::makeExampleMatchedDataSet(type = "GRanges")
x
```

This creates a GRanges object with random ranges and 3 metadata colums (`feature1`, `color`, and `length`) that describe "features" of each range.

Let's demonstrate generating a set of matched ranges by finding ranges that are `FALSE` for `feature1` but have the same distributions of `color` and `length`.

Our `focal` set of interest is where feature1 is `TRUE`, and `pool` is all other ranges (where feature1 is `FALSE`):

```{r, message=FALSE, warning=FALSE}
focal <- x[x$feature1]
pool <- x[!x$feature1]
```

We can use `nullranges::matchRanges()` to select a subset from pool that is distributionally matched in color and range length:

```{r, message=FALSE, warning=FALSE, results='hold'}
library(nullranges)
mgr <- matchRanges(focal = focal,
                   pool = pool,
                   covar = ~color + length,
                   method = 'stratified',
                   replace = FALSE)

mgr
```

`overview` can show how successful our matching was:

```{r}
overview(mgr)
```

## Visualizing with BentoBox

Defining the region to plot:

```{r, message=FALSE, warning=FALSE}
library(BentoBox)
region <- bbParams(chrom = "chr1", chromstart = 1, chromend = 1000)
```

We can visualize these ranges by creating a `BentoBox` page:

```{r, message=FALSE, warning=FALSE, fig.width=10, fig.height=10}
bbPageCreate(width = 8.5, height = 6.5, default.units = 'inches')
```

Visualize the `pool` GRanges:

```{r, message=FALSE, warning=FALSE, results = 'hold', fig.width=10, fig.height=10}
poolSet <- bbPlotRanges(data = pool,
                        params = region,
                        x = 1,
                        y = 1,
                        width = 2.5,
                        height = 2.5,
                        fill = pool$color)

bbAnnoGenomeLabel(plot = poolSet,
                  x = 1,
                  y = 3.55)

bbPlotText(label = "Pool Set",
           x = 2.25,
           y = 0.9,
           just = c("center", "bottom"),
           fontcolor = "#33A02C",
           fontface = "bold",
           fontfamily = 'mono')


```

Visualize the `focal` GRanges of interest:

```{r, message=FALSE, warning=FALSE, results = 'hold', fig.width=10, fig.height=10}
focalSet <- bbPlotRanges(data = focal,
                         params = region,
                         x = 5,
                         y = 1,
                         width = 2.5,
                         height = 1,
                         fill = focal$color)

bbAnnoGenomeLabel(plot = focalSet,
                  x = 5,
                  y = 2.05)

bbPlotText(label = "Focal Set",
           x = 6.25,
           y = 0.9,
           just = c("center", "bottom"),
           fontcolor = "#1F78B4",
           fontface = "bold",
           fontfamily = 'mono')
```

And finally the `matched` GRanges:

```{r, message=FALSE, warning=FALSE, results = 'hold', fig.width=10, fig.height=10}
## Matched set
matchedSet <- bbPlotRanges(data = mgr,
                           params = region,
                           x = 5,
                           y = 2.5,
                           width = 2.5,
                           height = 1,
                           fill = mgr$color)

bbAnnoGenomeLabel(plot = matchedSet,
                  x = 5,
                  y = 3.55)

bbPlotText(label = "Matched Set",
           x = 6.25,
           y = 2.75,
           just = c("center", "bottom"),
           fontcolor = "#A6CEE3",
           fontface = "bold",
           fontfamily = 'mono')
```

Below our GRanges, we can use `BentoBox` to place ggplots created from the `nullranges` `plotPropensity()` and `plotCovariate()` functions.

First lets plot propensity scores:

```{r, message=FALSE, warning=FALSE, results='hold', fig.width=10, fig.height=10}
library(ggplot2)
smallText <- theme(legend.title = element_text(size=8),
                   legend.text=element_text(size=8),
                   title = element_text(size=8),
                   axis.title.x = element_text(size=8),
                   axis.title.y = element_text(size=8))

## Plot propensity scores ggplot in BentoBox
propensityPlot <-
  plotPropensity(mgr, sets=c('f','m','p')) +
  smallText +
  theme(legend.key.size = unit(0.5, 'lines'),
        title = element_blank())

bbPlotGG(plot = propensityPlot,
          x = 1, y = 4.5, width = 2.5, height = 1.5,
          just = c("left", "top"))

## Text labels
bbPlotText(label = "plotPropensity()",
            x = 1.10, y = 4.24,
            just = c("left", "bottom"),
            fontface = "bold",
            fontfamily = 'mono')
bbPlotText(label = "~color + length",
            x = 1.25, y = 4.5,
            just = c("left", "bottom"),
            fontsize = 10,
            fontfamily = "mono")
```

plotting `color` and `length` covariates:

```{r, message=FALSE, warning=FALSE, results='hold', fig.width=10, fig.height=10}

## Plot "color" covariate
covColor <-
  plotCovariate(x=mgr, covar=covariates(mgr)[1], sets=c('f','m','p')) +
  smallText +
  theme(legend.text = element_blank(),
        legend.position = 'none')

bbPlotGG(plot = covColor,
          x = 3.50, y = 4.5, width = 1.8, height = 1.5,
          just = c("left", "top"))


## Plot "length" covariate
covLength <-
  plotCovariate(x=mgr, covar=covariates(mgr)[2], sets=c('f','m','p'))+
  smallText + 
  theme(legend.key.size = unit(0.5, 'lines'))

bbPlotGG(plot = covLength,
          x = 5.30, y = 4.5, width = 2.75, height = 1.5,
          just = c("left", "top"))


## Text labels
bbPlotText(label = "plotCovariate()",
           x = 3.75,
           y = 4.24,
           just = c("left", "bottom"),
           fontface = "bold",
           fontfamily = "mono")
bbPlotText(label = covariates(mgr),
           x = c(4, 5.9),
           y = 4.5,
           just = c("left", "bottom"),
           fontsize = 10,
           fontfamily = "mono")
```

When the plot is finished you can remove the page guides:

```{r, fig.width=10, fig.height=10}
bbPageGuideHide()
```

And now your plot is publication-ready!
